### 2022-01-03 writeup.tex

\documentclass[12pt, oneside]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath, amssymb, amsthm, latexsym, bm,enumerate, fancyhdr, tikz,afterpage, float}
\usepackage{graphics,graphicx,geometry}
\usepackage{hyperref,color}
%beamer uses these
\usepackage{etex}
\usepackage{epsfig, subfigure, psfrag}
%\usepackage[]{hyperref}
\usepackage[english]{babel}
\usepackage{multimedia}
\usepackage[all]{xypic}
\usepackage{alltt}
\usepackage[numbers,square,sort&compress]{natbib}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usepackage{xspace}
\usepackage{comment}
\usepackage{mathtools}

\usepackage[framemethod=tikz]{mdframed}
%color definitions

%berné added
\usepackage{pst-node}
%\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\colorlet{darkgreen}{green!50!black}
\definecolor{darkblue}{rgb}{0, 0, .6}
\definecolor{midnightblue}{rgb}{.2, .2, .7}
\definecolor{grey}{rgb}{.7, .7, .7}
\definecolor{Red}{rgb}{.9,0,0}
\definecolor{Blue}{rgb}{0,0,.9}
\definecolor{Green}{rgb}{0,.9,0}

\newenvironment{greenblock}[3]{%
  \setbeamercolor{block body}{bg=structure!10,fg=black}
  \setbeamercolor{block title}{bg=green!20,fg=black}
  \begin{block}{#1}}{\end{block}}

\newenvironment{redblock}[3]{%
  \setbeamercolor{block body}{bg=red!10,fg=black}
  \setbeamercolor{block title}{bg=red!20,fg=black}
  \begin{block}{#1}}{\end{block}}

%%% more color definitions

\colorlet{lightred}{red!30!white}
\colorlet{lightblue}{blue!30!white}
\colorlet{lightyellow}{yellow!30!white}
\colorlet{keylime}{green!10!white}
\colorlet{darkgreen}{green!50!black}

\def\docolor#1#2{\textcolor{#1}{#2}}
\def\highlight#1{\docolor{blue}{#1}}
\def\underscore#1{\underline{#1}}
\def\boldify#1{{\bfseries #1}}

\restylefloat{figure}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{nt}[thm]{Note}
\newtheorem{exs}[thm]{Examples}
\theoremstyle{definition}
\newtheorem{dfn}[thm]{Definition}
%\theoremstyle{remark}
\newtheorem*{ex}{Example}
\newtheorem{rmk}[thm]{Remark}
\newtheorem*{keg}{Key example}
\newtheorem*{keyeg}{Key example}
\newtheorem*{term}{Terminology}
\newtheorem*{notn}{Notation}
\newcommand{\mt}{}
\newcommand{\del}{\ensuremath{\backslash}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
% \newcommand{\beq}{\begin{eqnarray}}
% \newcommand{\eeq}{\end{eqnarray}}
% \newcommand{\beqn}{\begin{eqnarray*}}
% \newcommand{\eeqn}{\end{eqnarray*}}
\newcommand\ix[1]{#1\index{#1}}
\newcommand{\ls}{\leqslant}
\newcommand{\gs}{\geqslant}
\newcommand\lset[2]{\left\{\left.#1\ \right|\ #2\right\}}
\newcommand\rset[2]{\left\{#1\ \left|\ #2\right.\right\}}
\newcommand{\ol}{\overline}
\newcommand{\mbb}[1]{\ensuremath{\mathbb{#1}}}
\newcommand{\mbf}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\bbfq}{\mathbb{F}_q}
\newcommand\bbp{\mathbb{P}}
\newcommand{\bbc}{\mathbb{C}}
\newcommand{\bbf}{\mathbb{F}}
\newcommand{\bbq}{\mathbb{Q}}
\newcommand{\bbr}{\mathbb{R}}
\newcommand{\bba}{\mathbb{A}}
\newcommand{\mcal}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\CC}{\mcal{C}}
%\newcommand{\CD}{\mbox{$\cal D$}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\CE}{\mcal{E}}
\newcommand{\CF}{\mcal{F}}
\newcommand{\CG}{\mcal{G}}
\newcommand{\CO}{\mcal{O}}
%\newcommand\weight{\operatorname{weight}}
\newcommand{\wt}{\operatorname{weight}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\columnrank}{\operatorname{column~rank}}
\newcommand{\rowrank}{\operatorname{row~rank}}
\newcommand{\im}{\operatorname{Im}}
\newcommand{\ifff}{\Longleftrightarrow}
\newcommand{\ham}{\operatorname{Ham}}
\newcommand{\tran}{^\textup{T}}
\newcommand{\li}{linearly independent}
\newcommand{\ld}{linearly dependent}
\newcommand{\sgn}[1]{\ensuremath{\mathrm{sgn}(#1)}}
\newcommand{\krn}[1]{\ensuremath{\mathrm{Ker}(#1)}}
\newcommand{\ch}[1]{\ensuremath{\mathrm{char}(#1)}}
 \newcommand{\jmod}[2]{\mbox{ $\equiv #1$}\mbox{ \rm (mod $#2$)}}
\newcommand{\group}[2]{\ensuremath{(#1,\, #2)}}
\newcommand{\ring}[3]{\ensuremath{\langle #1,\, #2,\, #3 \rangle}}
\newcommand{\gl}[2]{\ensuremath{\mathrm{GL}(#1,\, #2)}}
%\newcommand\ex[1]{\emph{#1}\index{#1|btemp}}
\newenvironment{pfenum}{\begin{proof}\indent\begin{enumerate}\vspace{-\topsep}}{\qedhere\end{enumerate}\end{proof}}
\makeatletter
\def\imod#1{\allowbreak\mkern10mu({\operator@font mod}\,\,#1)}
\makeatother

\pagestyle{fancy}
\fancyhead[L]{Draft 1}
% \fancyhead[C]{Algebra}
\fancyhead[R]{\leftmark}

% \fancypagestyle{plain}{
% \fancyhead[L]{WTW 381}
% \fancyhead[C]{Algebra}
% \fancyhead[R]{\leftmark}}

\renewcommand{\sectionmark}[1]{\markboth{\thesection.\ #1}{}}

\renewcommand{\headrulewidth}{0.4pt}

\addtolength{\headheight}{\baselineskip}

\sloppy

\begin{document}

\title{Title}

\author{B. Nortier\\
% {\small \it Department of Mathematics \& Applied Mathematics, University of Pretoria, Pretoria 0028} \\ {\small Email: u17091820@tuks.co.za \& } \\
}
\date{2021}
\date{}
\maketitle

\begin{abstract}
Abstract goes here.
\end{abstract}

% {\bf Keywords} Data-driven dynamical systems, Koopman Operator, Causal Embedding, Nonautonomous Dynamical Systems.

% \newpage

\section{Introduction}
Traditional methods for modelling physical systems involves deriving governing equations by
considering the world and discover some fundamental relationships between encountered
quantities and phenomena. These fundamental relationships are then expressed mathematically and
used as the foundations to derive more complicated relationships. Consider the example of
Newton’s law F=ma
However our world around is, for the most part, much more complex than that which can be
distilled into elegant equations. Many of the more complex systems we don’t fully understand, nor
even have good physical intuition of.
To complicate this – many of the physical phenomena we observe in nature are chaotic.
Mathematically, this would imply what we call sensitive dependence on initial conditions. Despite
having highly similar initial conditions, differing orbits will diverge quickly and by such a large degree
that it becomes seemingly impossible to retrace them steps back to the origin conditions.The
presence of the smallest of computational noise or measurement error makes long-term pointwise
prediction very difficult to minimise.
We may however exploit the environment of data-overload we currently find ourselves in. This is
one where an incredible amount of information is being gathered and generated on a daily basis.
Alongside this we have also witnessed the advent of greater computing power and storage.
This project work focusees on constructing governing models for real-world systems in some sense
‘blindly’, by making use of the abundance of information we have at our disposal. An advantage of
this approach is the fact that we need only make a very few starting assumptions about the system
we choose to investigate, allowing the methodology it to be greatly generalised. In establishing 
topological conjugacy between underlying systems we can then, amongst others, establish long-
term accurate results.

By solidifying the mathematical underpinnings of our theory, we may guarantee having the ability to
construct models with predictive power ranging from molecular biology to neuroscience.
The goal of the methodology described here is to be able to predict the dynamics into the future by
constructing a model that will accurately mimic the dynamics of a system and hence also provide a
framework for future evolution of a dynamical system.\\

\emph{Additional to Intro?}\\
Examples in literature attempting to forecast data from such systems abound: One could model the
unobserved states as stochastic quantities. Or one could map the data to a higher-dimensional space
by means of delayed-time map of observations and then hope to see some of the the unobserved
characteristics of the data revealed.

\newpage

\section{Possible addition or Post-introduction:}

\emph{Expand on non-autonomous systems, their unique challenges and chaotic dynamics?}

Long-term consistency while modeling \textbf{chaotic} dynamical systems is an issue. When systems have 
sensitive dependence on initial conditions, negligible numerical errors multiply within a very short 
time gap to the point that it becomes unfeasible to track specific orbits.
We’ll focus on case where attractor is present.

\section{Setup and Learning Problem}

Consider a relatively simple learning problem:

Given ($u_{0}$, $u_{1}$, …, $u_{m}$), forecast um+1, um+2, …, given that $u_{n}$ is defined by relation  $u_{n+1} = Tu_{n}$.

However, T is unknown.

More involved learning problem: We have $\theta$(u0), $\theta$(u1), …, $\theta$(um) in an unknown dynamical system (U,T). 
Forecast $\theta$(um+1), $\theta$(um+2)

\textbf{Now consider a more refined scenario:}
\begin{enumerate}
  \item  a continuous time system with true state u in U (input space)
  \item Measured signal = $x_{n}$ (where: $x_{n}$ homeo x(u(tn)) + $\eta_{n}$ <-- noise 
\end{enumerate}

We have a scalar time-series.

We construct multi-dimensional observable

How? 
\begin{enumerate}
  \item Delay-coordinates: Given {x1,x2,$\ldots$}: define $x_{n}$ = [$x_{n}$,$x_{n}$-$\tau$,$\ldots$, $x_{n}$-(d-2)$\tau$,$x_{n}$-(d-1)$\tau$].T $\in\mathbb{R}^{d}$ 
  \item $\tau$=lag
  \item d=embedding dimension (park those two for now)
\end{enumerate}

Given: $x' = f(x)$ where $x(t)\in\mathbb{R}$  (x(t) are the states, f is assumed smooth)

Define flow: $\Phi$(x(t0),M)=x(t0+M), $\Phi$:U$\times$R $\rightarrow$ U

we can relate the two:

d/dt$\Phi$ (x(t0),t) = d/dt x(t0+M) = f(x(t0+M)) =f($\Phi$(x(t0),t)

Now define time-map $\Phi$M : x(t0+M)= $\Phi$T(x(t0))

i.e. x(t0+kM) = $\Phi$k

M(x(t0))= $\Phi$M$\circ$ $\Phi$M $\circ$$\ldots$$\circ$ $\Phi$M(x(t0) = k-fold composition

BUT: we only get to see a 1D time series $\theta$(x(t)), where $\theta$: R

N$\rightarrow$R is a smooth observation/measurement function.

Can information about the system state x(t) be retained in this time series data $\theta$(x(t))?

If we know f, then easily. But we don’t.

\textbf{However: Takens’ theorem gives some hope by saying yes! }

\subsection{Some side-questions}
\emph{Let’s talk about the fact that the manifold U is an attractor etc? Takens only holds for attractors, so
that is then why we immediately “restrict” U by only considering the subset containing the attractor
right? (ie. Only consider the space occupied by attractor. This is then why we demand that U is
compact and T surjective?
What about density of attractors? (1)}

$\ldots$

\section{Takens Theorem}
\begin{enumerate}
  \item What's that?
  \begin{enumerate}
        \item Takens defines delay-coord-map F$\theta$
        \begin{enumerate}
              \item F$\theta$ : [$\theta$(T-2du), $\ldots$ , $\theta$(T-1u), $\theta$(u)] $\rightarrow$ [$\theta$(T-2d+1u), $\ldots$ , $\theta$(T-1u), $\theta$(Tu)]
              \item Where T is flow T$u_{n-1}$ = $u_{n}$ (previously our flow was $\Phi$)
              \item $\theta$ is measurement function,
        \end{enumerate}  
        \item When confining DS to manifold U (in which system is contained), Takens showed
        that, if certain smoothness conditions are satisfied on T and $\theta$ then delay-coord
        map F$\theta$ is embedding of U onto the reconstruction space for \emph{almost every
        choice(2)} of measurement function $\theta$ (the observable)
        \item Or alternatively: Takens' embedding theorem guarantees that almost all
        dynamical systems can be reconstructed from just one noiseless observation
        sequence.
        \item i.e. for great number of possible observation functions $\theta$, F$\theta$ preserves the
        topology of U
        \item \emph {See Graph 1 }
        \begin{enumerate}
          \item If we are in a state u, and then evolve towards state Tu before  embedding into R(2m+1), then it is the same as embedding into R(2m+1) and then evolving through the map F$\theta$
        \end{enumerate}
        \item Thus: information about U can be retained in the time series’ (observation)
        output. By preserving the topology of the manifold U in the reconstruction space
        X, topological invariants of manifold are preserved, of which dimensionality is
        one (again considered later important later. (3)
  \end{enumerate}
  \item Problems with Takens
  \begin{enumerate}
    \item Even if we can find F$\theta$, our approximation of F$\theta$ is a map from a larger set R2m+1
    containing the embedded attractor. No theoretical guarantees that F$\theta$ will keep
    this attractor the same. (should have defined attractor earlier?)
    \item Takens only holds for noiseless observations.
        \begin{enumerate}
          \item Assume F was learnt but due to measurement error/noise we get V, not an attractor, then will tend somewhere else. This problem will later be removed by some properties of g (USP/UAP)
        \end{enumerate}
    \item Takens only holds if T is invertible. 
    \item  Functional complexity and stability of embedding F$\theta$ depend on choice of  observable $\theta$
  \end{enumerate}
\end{enumerate}

Now: Although we cannot necessarily find a map T:U $\rightarrow$ U, there does exist a map from
$\Phi$: $\Phi$2d,f(um) $\rightarrow$ $\Phi$2d,f(um+1) whenever U is a manifold of dimension less than d (should we say more
about d?)

\emph{Graph 2}
(In sketch: W==U )

\textbf{We’re one step closer to recovering information about the system states}

\section{Problem}

We're trying to approximate a map that might not even exist.

\emph{Goal:} 
predict dynamics into future and construct model from data which mimics dynamics.

\section{So how do we guarantee existence?}

\begin{enumerate}
  \item Assumptions
  \begin{enumerate}
      \item assume a compact, discrete time, dynamical system (U,T) (implicitly T cts)
      \begin{enumerate}
          \item T must be surjective. (This is important, we return to it later)  
      \end{enumerate}
      \item  A driven dynamical system g, who takes input in U. (implicitly g is cts)
      \begin{enumerate}
         \item g will be randomly initiated neural net <--  wil explain why later
        \item  g($u_{n}$-1, $x_{n}$-1) = $x_{n}$
      \end{enumerate}
      \item Bi-infinite sequence u <-- physically: running for long time
  \end{enumerate}

  \item We can easily show existence of the function g:
  \begin{enumerate}
    \item   g = identity on X 
    \newline But then: g ignores the input u and any constant bi-infinite sequence would be a solution to input u.
    \item  However this gives trivial soln, can we find something better?
    \item  Define: reachable soln space XU, entire solutions, YT
    \begin{enumerate}
      \item Give own example of entire sol'n
      \item \emph{Graph 3}
    \end{enumerate}
  \end{enumerate}
  \item Choosing g
  \begin{enumerate}
      \item Choosing g st “not quench temporal structure in u”
      \item(4)
      \item  g must be cts, invertible (for SI-invertibility)
      \item Specific w.r.t tanh:
       \newline \emph{Describe set of all solutions of g for given input u}
  \end{enumerate}
\end{enumerate}

\section{Can we show uniqueness of an entire solution?}

\emph{Graph 4}

\begin{enumerate}
  \item UAP: regardless of starting position, all trajectories converge to a single trajectory. This
  trajectory is the unique solution sequence x to the input sequence u from the previous
  slide.
  \item  USP: \emph{insert more info here}
  \begin{enumerate}
      \item The USP - we can start the system anywhere and in time it will forget
      \item  No new complexity
      \item USP $\equiv$ g is topological contraction $\equiv$ UAP (we’ll show this) 
      \begin{enumerate}
            \item 1. Define topological contraction      
      \end{enumerate}
  \end{enumerate}
  \item but infinite sequences are difficult to manage
  \begin{enumerate}
      \item   i. If both SI-invertibility and USP hold for g, then our causal mapping becomes much simpler: we can drop the infinite sequence to focus only on the most recent two terms. (Call it H2)
      \newline \textbf{AND} H2 becomes embedding    
  \end{enumerate}
\end{enumerate}

\section{Establishing Embedding and Conjugacy}

\begin{enumerate}
        \item Define: Embedding, Causal Embedding, Topological Conjugacy
        \begin{enumerate}
            \item Explain conjugacy more perhaps?  
        \end{enumerate}
        \item \emph{Graph 5}
        \item SI-invertibility, guaranteeing us the function $G_{T}$ on X$\times$X. <--- theoretical basis to try approximate!
        \item If g is SI-invertible and {$u_{n}$} in U is an orbit of T then:
        \begin{enumerate}
          \item   $G_{T}$ exists
          \item if g also has USP then (YT, $G_{T}$) top.semi-conjug to ($\hat{U}$, $\hat{T}$)
          \item if T:U $\rightarrow$ U is homeo then (YT, $G_{T}$) is top.conjug to ($\hat{U}$, $\hat{T}$) and top.conjug to (U,T) so g is causal embedding
          \item WE can learn single-delay dynamics
          \begin{enumerate}
            \item forecast future values of $G_{T}$
            \item forecast future values of $u_{n}$
          \end{enumerate}
        \end{enumerate}
        \item Question: why can one learn single-delay lag dynamics of the driven states through $G_{T}$
        with \textbf{enough} data?
\end{enumerate}

\section{Why $G_{T}$?}

Why do we do this?

\begin{enumerate}
  \item Essentially embed attractor into higher dimensional space X$\times$X.
  \item  Hence: more `dimensional room' for the dynamics to move, and we might hope
  that the dynamics of $G_{T}$ is in some sense simpler than that of T.
  \newline \textbf{OR} 
  \newline  Embed into state space with much higher dimension s.t. unobserved features of sequential iput data ($u_{n}$) can emerge in ($x_{n}$)
\end{enumerate}

\section{Progression to $\Gamma$ma instead of $G_{T}$}

\begin{enumerate}
  \item Define $\Gamma$:($x_{n}$-1, $x_{n}$) $\rightarrow$ $u_{n}$
  \begin{enumerate}
      \item Existence? $\Gamma$ exists when G exists  
  \end{enumerate}
  \item New conjugacy  
  \newline \emph{Graph 6}
  \item Why not $G_{T}$?
  \begin{enumerate}
    \item dimension on which $G_{T}$ is defined can be high (wasn’t it the goal to increase dimension?)
    \item  If we know $\Gamma$ then we can drive the system autonomously. 
    \newline  And then we know $G_{T}$ anyway
    \item Also: numerical errors could accumulate  when working
    \newline \emph{Graph 7}
  \end{enumerate}
\end{enumerate}

\section{Programmatically} 
\begin{enumerate}
  \item Implementation
  \begin{enumerate}
    \item Explain the choice of g
    \item Explain DC, SDD, MDD, AMDD
    \item Sparse matrices
    \item Explain why RNN is chosen (above, for example, FNN)
  \end{enumerate}
  \item Some Examples
  \begin{enumerate}
    \item Double Pendulum successes
    \item Data dying down (ie. Not surjective and make more remarks here on surjectivity as promised in the beginning)
    \item Attractors from previous article such as Lorenz, Henon
    \item Other attractors I’ve experimented with.
  \end{enumerate}
\end{enumerate}

\section{Comparison with/Mention of previous work}

\begin{enumerate}
  \item Mention previous work, such as SyndiPi, and that it doesn’t work
  \item  Data-driven approaches alongside machine-learning algorithms have outperformed the classic delay-coordinate embedding as formulated by Takens in the prediction of chaotic dynamical systems’ evolution.
\end{enumerate}

\section{Further Work}

MD

\section{More things to Add}
\emph{Unsure here whether it's important}

\begin{enumerate}
  \item Defining attractors
  \item Instability due to perturbation and how prevented
  \item Defining a process?
  \item \emph{Graph 8}
  \item State contractions and their importance
  \item  More actual theorems
  \item Worry:
  \begin{enumerate}
    \item  So far I have very little “mathematical” theorems or proofs. Is this a problem
    \newline This will, however, probably be added when expanding on attractors and USP
  \end{enumerate}
  \item Should I show/state that USP == UAP (equivalent)?

\end{enumerate}

% \section{Definitions}

% \begin{dfn}{Discrete-time autonomous system}
%   A discrete-time autonomous system on a state space $X$ is given by the map $g:X\mapsto{X}$, where the dynamics are generated by compositions of $g$ with itself.
% \end{dfn}

% The system evolves according to the equation $x_{n} = g(x_{n-1})$ for $n\geq1$ and generates the sequence
% $\ldots, x_{-1}, x_{0}, x_{1},\ldots$.

% \begin{dfn}{Discrete-time nonautonomous systems}
%     A discrete-time nonautonomous system on a state space $X$ is then a collection of continuous maps ${g_n}$ such that $g_n:X\mapsto{X}$ satisfying the equation
%     $x_{n} = g_{n-1}(x_{n-1})$ at each timestep $n$.
% \end{dfn}

% Now consider a subset of this discretised version of nonautonomous systems in which external factors are taken into account.

% \begin{dfn}{Input-driven system}
%   An input-driven system is understood to be a nonautonomous, discretised dynamical system on the state space $X$ consisting of an input, or driving, sequence, 
%   ${u_{n}}\subset{U}$ and a continuous function $g:U\times{X}\mapsto{X}$ such that at time $n$ the state $x_{n}$ is determined by $x_{n}=g(u_{n-1}, x_{n-1})$
% \end{dfn}

% \begin{rmk}
%   $g^{n}$ denotes the self-composition of the function $g:U\times{X}\mapsto{X}$ n times.  
% \end{rmk}

% \begin{dfn}
% Discrete-time state space models of the form 
% \begin{equation}
%     x_{n+1} = g(u_{n}, x_{n})
% \end{equation}
% \begin{equation}
%     y_{n+1} = f(x_{n+1}),
% \end{equation}
% where $n \in \mathbb{Z}, u_{n}$ belongs to an input space $U$ and state $x_{n}$ belonging to a state space $X, g: U \times X \rightarrow X$, and a measurement $f:X \rightarrow Y$
% \end{dfn}

% \begin{dfn}
% The dynamics on $X$ are generated by the update equation $x_{n+1} = g(u_{n},x_{n}$
% \end{dfn}

% \begin{dfn} {Entire solution} \\
% Given a driven system $g$ and an input $\overline{u}$, we call a sequence $\{x_{n}\}_{n\in\mathbb{Z}})$ an entire-solution, or just a solution, if it satisfies the update equation $x_{n+1} = g(u_{n},x_{n}$ for all $n\in\mathbb{Z}$
% \end{dfn}

% \begin{dfn} {Entire solution of process a $\phi$} \\
%   An entire solution of a process $\phi$ on the state space $X$ is a sequence $\{x_{n}\}_{n\in\mathbb{Z}}$ where $x_{n}\in{X}$ for all $n\in\mathbb{Z}$ and 
%   $\phi(n,m,x_{n})=x_{n}$ whenever $n\leq{m}$.
% \end{dfn}

% \begin{dfn} {$\phi$-Invariant Set}
%   The set $\mathbb{A} = \{A_{n} : A_{n} \subset X\}$ where $\mathbb{A}_{m}\subset X$ for all $m$ and $\phi(n,m,A_{m})=A_{n}$ for $n\geq{m}$.
% \end{dfn}

% \begin{rmk}
%   Note: when we speak of g, then we refer implicitly to the entire underlying dynamical system.
% \end{rmk}

% The infinite product space $\overline{Y}$ is defined as follows: $\overline{Y}:=\prod_{i=-\infty}^{\infty}{Y_{i}}$ such 
% that $Y_{i} \equiv$.

% \begin{dfn} {Unique Solution Property} \\
% A driven system $g$ has the unique solution property (USP) if for each input $\overline{u}$ there exists exactly one solution. \\
% Expressed more rigourously, we may state that $g$ has the USP if there exists a well-defined map $\Psi:\overline{U}\mapsto\overline{X}$.
% $\Psi(\overline{u})$ then represents the unique entire solution when the input is $\overline{u}$. 
% % Alternatively, g has the USP if there exist a well-defined map $\Psi$ between the infinite product space $\overline{U}$ and $\overline{X}$, where $\overline{Y}:=\prod_{i=-\infty}^{\infty} Y_{i}$ and $Y_{i}\equivY$ so that $\Psi(\overline{u})$ denotes the unique solution obtained from the input $\overline{u}$.

% \end{dfn}

% \begin{dfn}
% Echo State Property or Topological Contraction
% \end{dfn}

% \begin{thm}{Contraction Mapping Theorem}
%   $f:M\rightarrow M$ where $(M,d)$ is a metric space then $\exists k\in[0,1]$ such that (s.t.) $\forall{x,y}\in\mathbb{R}$ we have
%   $d(f(x),f(y)) \leq d(x,y)$
% \end{thm}

% \begin{dfn}
% Dissipative System
% \end{dfn}

% \begin{dfn} {Basin of Attraction }
% Basin of Attraction for a stable fixed point
% \end{dfn}

% \begin{dfn}
% Local Irreducible Attractor
% \end{dfn}

% \begin{dfn}
% Pullback Attractor
% \end{dfn}

% \begin{dfn}
% Forward Attractor
% \end{dfn}

% \begin{dfn}
% Uniform Attraction Property:
% $\forall \epsilon \exists N \in \mathbb{N} s.t. \forall y_{0} \in X \forall \overline{u}\in\overline{U}, d(x_{n}, y_{n})<\epsilon holds for n>\mathbb{N}.$ Here $y_{n+1}:=g(u_{n},y_{n})$ for $n\geq0$ and $\{x_{n}\}$ is a solution for input $\overline{u}$
% \end{dfn}

% \begin{dfn}
%     $\hat{U}_{T} = $ the space of all left-infinite T-orbits = $\{(\ldots, u_{-2}, u_{-1} : Tu_{n} = u_{n+1})\}$
% \end{dfn}

% \begin{dfn}
% Left-Infinite Sequence Space $\overleftarrow{U}$ where

% $\overline{Y}:=\prod_{i=-\infty}^{\infty} Y_{i}$ and
% $Y_{i} \equiv Y$ equipped with the product topology.

% \end{dfn}

% \begin{dfn}
% The left-infinite fragment of any bi-infinite input sequence $\overline{u}$ is denoted by $ \overleftarrow {u}^{n}:=($\ldots$,u_{n-2}, u_{n-1})$
% \end{dfn}

% \begin{dfn}
% The reachable set of a driven system ${g}$ is the union of all elements of all possible solutions. \\
% $X_{U} := \{ x \in X : \exists$ a solution $\{x_{n}\}$ for some $\overline{u}$ and $x=x_{k}$ for some $k \in \mathbb{Z}$
% \end{dfn}

% \begin{rmk}
% An input value is denoted as $v$.
% \end{rmk}

% \begin{dfn}
%   Orbit of T
% \end{dfn}

% \begin{dfn}
%   \begin{equation}
%     Y_{T} = \{ (x_{n-1}, x_{n}) : \{x_{k}\}  \textrm{is a solution of some orbit of} T \}  
%   \end{equation}
% \end{dfn}

% \begin{rmk}
%   The following notation is used:\\
%   $\overleftarrow{u} = (\ldots, u_{-2}, u_{-1})$ \newline
%   $\overleftarrow{u}^{n} = (\ldots, u_{n-2}, u_{n-1})$
% \end{rmk}

% % \begin{def}
% % The update of the input is defined as $u_n$. 
% % \end{def}

% \begin{dfn}
% The input up to time $n$ with an input $v$ at time $n$ is denoted 
% $\overleftarrow{u}^{n}v:=(\dots,u_{n-2}, u_{n-1}, v)$.
% \end{dfn}

% \begin{dfn}
% H is a causal mapping if $H:\overleftarrow{U}\rightarrow\overleftarrow{X}_{U}$ is a continuous and surjective map such that 
% \end{dfn}

% \begin{dfn}{Topological Conjugacy}
% A homeomorphism between two dynamical systems (U,T) and (V,S) is a a function $\phi:U\rightarrow V$ such that $\phi\circ\ T = S\circ\phi$.\\ We then say that 
% (V,S) is \emph{topologically conjugate} to (U,T).
% \end{dfn}

% When $\phi$ is merely continuous and not a homeomorphism, we call the map a topological semi-conjugacy. This many-to-one relationship guarantees at least a 
% coarse-grained depiction of the original system.

% \begin{dfn}
%   Universal Semi-Conjugacy: A map $h:\overleftarrow{X}\rightarrow X_{U}$ which is continuous and surjective s.t. $\forall v\in U$ the diagram commutes: \\
%   \textbf{XXX}
%   \end{dfn}

% \begin{dfn}{Embedding}
% A homeomorphism $f:Z\rightarrow Y$ for $Y\in X$. \\
% We then say that $Z$ is embedded in $X$ by $f$
% \end{dfn}

% \begin{dfn}{Observable}
% A one-dimensional time-series $y(t)=\phi(x(t))$, where $\phi:\mathbb{R}^{n}\rightarrow\mathbb{R}$ is a smooth so-called observation or measurement function.
% \end{dfn}

% \begin{dfn} {Process}\\
%   Consider the subspace $\mathbb{Z}_{\leq}^{2} := {(n,m) : n,m \in \mathbb{Z}, n\geq{m}}$. \\
%   The process $\phi$ is a mapping $\phi:\mathbb{Z}_{\leq}^{2}\times{X}\mapsto{X}$ satisfying the following:
%   \begin{enumerate}
%     \item $\phi(m,m,x)=x $ \quad for all $ m\in\mathbb{Z}, x\in{X}$
%     \item  $\phi(n,m,x)= \phi(n,k,\phi(k,m,x)) \quad $ for all $ m,k,n \in\mathbb{Z}$ where $m\leq{k}\leq{n}$, $x\in{X}$
%     \item $\phi(n,m,\dot)$ is continuous on $X \forall{n,m}$ 
%   \end{enumerate}
% \end{dfn}

% % Process: Two-parameter semigroup formalism (10,15) defined for nonautonomous systems

% \begin{thm}
%   A nonautonomous system $g_{n}$ on $X$ exists if and only if a process $\phi$ on $X$ exists.
% \end{thm}

% The proof is almost trivial:
% \begin{proof}
%   A process $\phi$ on $X$ generates a nonautonomous system $\{g_{n}\}$ if we define $g_{n}(x):=\phi(n+1,n,x)$. In so doing we obtain the 
%   system state $x_{n}$ at time $n$ by $\phi(n,m,x)$ where $x_{0}$ was the initial state at time $m$.

%   Conversely, the nonautonomous system  with $\{g_{n}\}$ generates a process $\phi$ on $X$ if  we define as follows: 
%   $\phi(m,m,x):=x$ and $\phi(n,m,x):= g_{n-1}\circ{\ldots}\circ{g_{m}(x)} = x_{n}$.   
% \end{proof}

% % the nonautonomous system $(X, g)$

% \begin{dfn} {Flow} \\
%   The unique global solution $\Phi(t,x_{0})$  of $x'=f(x)$ passing through some $x_{0}$ at $t=0$ where
%    $\Phi$ satisfies the following properties:
%    \begin{enumerate}
%     \item $\phi(0,x_0)=x_{0}$
%     \item $\phi(t,x_{0})$ is continuous on its domain
%     \item $\Phi(t+s,x_{0})=\Phi(t,\Phi(s,x_{0}))$
%    \end{enumerate}
% \end{dfn}

% It follows that $\Phi(t,x_{0})$ attains the value $x_t$ at time $t$, and so we may define the function 
% $T(x_0):=\Phi(M,x_{0})$ , which can be referred to as the \emph{Time-T Map}. 
% \newline

% Now consider the following: \\
% \begin{equation}
%   T\circ{T(x_0)} = T\circ{\Phi(M,x_{0})} = \Phi(M,\Phi(M,x_{0})) = \Phi(2M,x_{0}) = T^{2}(x_0)
% \end{equation}

% In general then: {  $T^{n}(x_{0}) = $ the value of the solution after  nxM  time-units} and any solution $(x_k)$in U is of 
% the form ${x, Tx, T^{2}x, \ldots}$, i.e. it satisfies the difference equation 
% \begin{equation}
%   x_{k+1} = T(x_k)
% \end{equation}

% \begin{dfn}{Dynamical System} \\
% A tuple $(U,T)$ for an attractor/state space U which is a compact metric space and a function T that is not necessarily continuous.
% \end{dfn}

% It may easily be verified that if T is obtained from a flow, then T is a homeomorphism. \emph{CITATION}\\

% One can also assume, without loss of generality, that T is surjective. To see this, consider 
% $\ldots$  \emph{CITATION}

% The evolution of a dynamical system may happen smoothly over time or over discrete time-units. Though evolution in nature occurs
% in a continuous fashion, we may only observe this change in discrete-time due to the fact that we can only gather data through
% observations at specific intervals of time. It is thus necessary to discretise the system. \emph{(which automatically arise per dis-
% cretization in simulations on digital computers).}

% \bigskip
% Assume that the system has state $u$ in the input space $X$. We cannot directly observe u, (for if we did then this project would 
% be superfluous). Rather, we have some observed signal $(x_{n})$ of the form
% \begin{equation}
% x_{n} = x(u)(t_n) + \nu_{n}
% \end{equation}
% for measurement noise $\nu{n}$.

% Measurement noise can be a great hindrance to accurate reconstruction of the system and could even become indistinguishable from a random time series. 
% \emph{The case of low-D data from high-D system}

% \subsection{Functions}

% \begin{equation}
%     \Gamma: Y_{T} \mapsto {U} \textrm{ s.t. } \Gamma:(x_{n-1}, x_{n}) \mapsto u_{n}
% \end{equation}

% \begin{equation}
%   \hat{T}(\overleftarrow{u_n}) = \hat{T}(\ldots, u_{-2}, u_{-1})
%   % & = (\ldots, u_{-2}, u_{-1}, Tu_{-1}) 
%   % & = (\ldots, u_{-2}, u_{-1}, u_{0})
% \end{equation}

% \bigskip
% \newpage

% \bigskip

\end{document}